---
title: "Exploratory"
author: "Bradley Paye"
date: "2025-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Motivation

A large literature studies price bubbles in the context of experimental asset markets. It is well known that such markets frequently produce bubbles in the sense of prolonged, upward deviations of asset prices from (known) fundamental value. However, theoretical models of rational bubbles imply specific forms of dynamics that amount to explosive price growth, or ``explosiveness.'' Although studies based on experimental asset markets typically report summary measures that plausibly suggest this form of price dynamics, we are aware of few studies that explicit test for explosiveness. In this paper, we apply time series econometric methods designed to test for the presence of explosive dynamics in prices. 

### R Bubbles Testing Package

We will use the R package "Multiple Bubbles" that incorporates many of the key  proposed by Phillips and coauthors in a set of papers.

In the following, we load up the package and do some very basic stuff.

```{r basic}
library(MultipleBubbles)

#simulate a set of random walks with 1/n drift

#set parameters
nseries <- 10
nsize <- 100
plotn <- 3
#simulate RWs
rw <- DGP(nsize,nseries)

#plot first few simulations

#plot(1:nsize,rw[,4],type = 'l')

# Plot multiple lines using matplot
matplot(1:nsize, rw[,1:3], type = "l", lty = 1, 
        col = c("red", "blue", "green"), xlab = "time", 
        ylab = "Y", main = "Simulated Random Walks")
legend("topright", legend = c("RW 1", "RW 2", "RW 3"), 
       col = c("red", "blue", "green"), 
       lty = 1)




```

## Including Plots

You can also embed plots, for example:

```{r spplot}
#Plot S&P data used in Phillips paper
plot(sp_data,type='l')
```

## Apply the basic ADF test to simulated random walks
To get started with the code package, we will explore the various elements, working our way up to the ultimate tests for explosiveness. The foundation for these more advanced tests is the classic augmented Dickey-Fuller or "ADF" test. The null hypothesis for the test is that the time series contains a unit root. In standard implementations, the alternative is that the time series is *stationary*. (The alternative in the tests for explosiveness will differ from this.)

The test can be formulated in a regression context, where the mode is:
$$
y_{t} = \alpha + \rho y_{t-1} + \nu_{t},
$$
in which $\nu_{t} \sim N(0,\sigma^{2}_{\nu})$.  The null hypothesis is $\rho = 1$ (unit root process) and the alternative is $|\rho| < 1$ (stationary process). One can estimate $\rho$ using OLS and construct a standard $t$-statistic; however, the issue is that under the null $y_{t}$ is nonstationary and this implies that, even under the normality assumption the statistic $t_{\rho}$ does not have a *t* distribution. 

It is common to transform the previous equation as follows in order to convert the test to one in which the null is that the slope parameter equals zero:

$$
\Delta y_{t} = y_{t} - y_{t-1} = \alpha + \beta y_{t-1} + \nu_{t},
$$
where we have the relation $\beta = \rho - 1$ and now the null is that $\beta = 0$ and the alternative $\beta < 0$. 



```{r adftest}
y <- rw[,1]
yy <- y[2:100]
xx <- y[1:99]
dy <- yy-xx

#Dickey-fuller by hand
df_reg <- lm(dy ~ xx)
summary(df_reg)

#set adf lags and model type
testlag <- 0
testmlag <- 5
modtype <- 1 #w/ constant and w/o trend

#This confirms that ADF w/ zero lags = DF
test_adf <- ADF_FL(y, adflag = testlag, mflag = modtype)
test_adf

test2_adf <- ADF_IC(y, adflag=testmlag, mflag=modtype, IC = 2)
test2_adf

#backward ADF crit values
foo <- badf(m = 100, t = 100, adflag = testlag, mflag = modtype)
```
```{r more adf}
library(tseries)
tryit <- adf.test(y,alternative = "s",k=testlag)
tryit

#reconcile with above by showing that test includes trend
modtype <- 2
test_adf <- ADF_FL(y, adflag = testlag, mflag = modtype)
test_adf

```
## Algorithm Summary
The algorithm proceeds as follows:

1. Obtain the minimal window length as integer part of $(0.01 + 1.8/\sqrt{T})T$. Call this $T^{\ast}$. Example: $T=100$, we have $0.19(T) = 19$ observations.
2. Run ADF regressions recursively forward starting with the initial $T^{\ast}$ observations. Retain the resulting ADF statistics. *Notea*: The Multiple Bubbles package function applies ADF tests with lag length determined by an information criterion set by user to either AIC or BIC. The model type (const,const and trend, no constant) must be set by the user, as well as the max lag. It seems there are no defaults so these must be explicitly set for the sadf_gadf function.
3. The SADF statistic equals the max of the recursive ADF statistics from the previous step. Inference is based on simulated critial values. 
```{r do sadf_gadf}

adflag_opt = 4  #max lag set to 4
mflag_opt = 1  #this is constant model
ICopt = 1  #1 = AIC
sadf_out <- sadf_gadf(y,adflag_opt,mflag_opt,ICopt)
sadf_out

```
